Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-02-07 Writing/appendix.rmd
made significant modifications to the manuscript and response to reviewers

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: Shunkei Kakimoto^[Corresponding author, Department of Agricultural Economics, University of Nebraska-Lincoln 102 Filley Hall 1625 Arbor Drive Lincoln, NE 68583, USA, E-mail address; skakimoto3@hunskers.unl.edu], Taro Mieno^[Department of Agricultural Economics, University of Nebraska-Lincoln, Lincoln, NE 68583, USA], Takashi S. T. Tanaka^[Artificial Intelligence Advanced Research Center, Faculty of Applied Biological Sciences, Gifu University, Gifu 501-1193, Japan], and David S. Bullock^[Department of Agricultural and Consumer Economics, University of Illinois, Urbana, IL 61801, USA]
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    global_numbering: true
    # reference_docx: word_template.docx
    reference_docx: word_template_format.docx
    # pandoc_args: ["-Fpandoc-crossref"]
bibliography: ML_VRA.bib
csl: computers-and-electronics-in-agriculture.csl
eqnPrefix: 
      - "Eq"
abstract: "Estimating site-specific crop yield response to changes in input (e.g., seed, fertilizer) management is a critical step in making economically optimal site-specific input recommendations. Past studies have attempted to estimate yield response functions using various Machine Learning (ML) methods, including the Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN) methods. This study proposes use of the Causal Forest (CF) model, which is one of the emerging ML methods that comprise \"Causal Machine Learning.\" Unlike previous yield-prediction-oriented ML methods, CF focuses strictly on estimating heterogeneous treatment effects (changes in yields that result from changes in input application rates) of inputs. We report results of using Monte Carlo simulations assuming various production scenarios to test the effectiveness of CF in estimating site-specific economically optimal nitrogen rates (EONRs), comparing CF with the yield-prediction-oriented ML methods RF, BRF, and CNN. CF's estimations of site-specific EONRs were superior under all scenarios considered. Furthermore, we also show that the quality of a model's yield predictions provides little if any information about the quality of its EONR predictions."  
---

```{r echo = F, include = FALSE}
library(knitr)
library(here)
here::i_am("GitControlled/Writing/manuscript_cea.rmd")
# opts_knit$set(root.dir = here())

options(htmltools.dir.version = FALSE)
options(knitr.duplicate.label = "allow")

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = FALSE, # <-- T
  cache.lazy = FALSE,
  echo = F
)
```

```{r, cache = FALSE}
# === packages ===#
library(tidyverse)
library(data.table)

library(png)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(raster)
library(bookdown)
library(png)
library(grid)
```

```{r analyses, results = "hide", fig.show='hide', cache = FALSE}
# /*=================================================*/
#' # Create figures and tables
# /*=================================================*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl(
  here("GitControlled/Codes/PrepareResults.rmd"),
  output = here("GitControlled/Codes/PrepareResults.R")
)

# --- Read the simulation results and prepare figures and tables ---#
source(here("GitControlled/Codes/PrepareResults.R"))
```

# Keywords: {-}

site-specific input management, nitrogen rate, on-farm precision experimentation, machine learning, causal forest, economically optimal input rates

# Introduction

Information about crop yield response input management is critical for the identification of site-specific economically optimal input rates (EOIRs) [@bullock1998does], and the difficulty in estimating site-specific yield response functions accurately within a field is the primary explanation of the persistent relatively low the adoption rate of variable rate crop input application technology [@lowenberg2019setting]. Historically, site-specific yield response functions have been estimated using the data generated from strip trials conducted on only parts of fields [@makowski2001statistical; @scharf2002corn; @bullock2002adding; @mamo2003spatial; @hurley2004estimating; @anselin2004spatial; @scharf2005field; @ruffo2006site; @Miao2006; @puntel2019development].
But recently whole-field on-farm precision experimentation (OFPE) [@bullock2019] has allowed implementation of improved statistical experimental designs (e.g., Latin Square, Randomized Block) on entire fields [@laurent_kyveryga_makowski_miguez_2019; @licht_witt_2019]. Data generated from whole-field OFPEs typically have greater local variation in experimental input application rates and soil/field characteristics than do strip trial designs, which has improved the statistical identification of heterogeneous impacts of input management changes on yields. The resultant availability of more spatially detailed experiment data has encouraged a resurgence of interest in estimating site-specific yield response functions. For example, @barbosa2020modeling, @krause2020random and @Gardner21 used OFPE data to estimate site-specific yield response functions and economically optimal site-specific N fertilizer and/or seed rates. These recent studies used machine learning (ML) methods extensively in place of more traditional parametric modeling approaches. @barbosa2020modeling applied the Convolutional Neural Network approach, @krause2020random used Random Forest-based approaches, and @Gardner21 used the shape-constrained generalized additive model. 
But this trend of applying ML methods may raises concerns. Historically, agronomists have worked meticulously to develop clever experimental designs to identify causes of yield response to input management. But the aforementioned recent applications of ML methodologies pay little attention to identify such causes, instead attempting to validate their models by estimating yield prediction accuracy. But the ultimate goal of OFPE is the accurate prediction of EOIRs, not yields. These are distinct objectives, and achievement of one need not imply the other.

The lack of and need for causal interpretation of the results of ML research applications have been increasingly recognized, in particular in economics and medicine [@athey2017beyond; @athey2018impact; @Athey2019a; @moraffah2020causal; @Storm2020; @arti2020research; @scholkopf2021toward], and causal machine learning (CML) methods were recently developed in response. Unlike traditional prediction-oriented ML methods, CML focuses on identifying causal impacts of an event (in our context, of an increase or decrease in the input application rate). The growing academic literature applying CML methods includes @richens2020improving, who's medical research pointed out that the inability of the traditional ML methods to identify cause-and-effect relationships between disease and patients' characteristics could lead to misdiagnoses, and showed that CML-based diagnoses could improve the accuracy of diagnostic prediction. In addition, @carbo2020machine used a CML to identify characteristics causing bank customers to adopt online banking services, and @bozorgi2020process applied CML to study the mining processes.

In this paper, we use the Causal Forest (CF) approach developed by @athey2016recursive and @Wager2018a to examine the accuracy of methods that estimate site-specific economically optimal N fertilizer rates (EONRs). We compare the EONR prediction performances of our CF-based approach to the widely used ML methods Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN), and we examine how yield prediction performance is related to EONR prediction performance, which allow us to examine whether model selection should be based on yield prediction accuracy when the ultimate goal is EOIR identifying. We use Monte Carlo simulations in which we synthetically generate yield and soil/field characteristics datasets to create stylized "field." Because real-world "true" EOIRs are unobservable, real-world data cannot be used to test the accuracy of statistical methods estimating EOIRs. This important difference contrasts starkly with yield prediction, since actual yields can be ground-truthed.

# Predicting Yield vs. Predicting Changes in Yield in Response to Changes in Input Management

In the following, we describe the fundamental information necessary to identify EOIR in general: the site-specific treatment effect of input application on yield. We then discuss the causal forest method in detail, show how it is particularly suitable to obtain this fundamental information, and contrast it to more traditional prediction-oriented ML methods, such as RF, BRF, and CNN.

## EOIR Estimation and Treatment Effects

Let $p$ be the price that a producer receives for a unit of a crop, $w$ be the per-unit price paid for a unit of an input, $OC$ represent other costs of production, and $y=f(I, \mathbf{c})$ describe the relationship between crop yield, the input application rate $I$ and a vector of field characteristics variables $\mathbf{c}$. A simple model of producer behavior assumes that the producer chooses a level of $I$ to maximize profits (revenues minus costs):

\begin{equation}
\max_{I} \quad pf(I, \mathbf{c}) - wI -OC. (\#eq:eqn1)
\end{equation}

Elementary calculus shows that the necessary condition for maximization is, that $\partial f(I, \mathbf{c})/\partial I - w =0$, which implicitly defines the economically optimal input application rate as a function of prices and soil/field characteristics, and may be arranged as \@ref(eq:eqn2):

\begin{equation}
\frac{\partial f(I^{*}(p,w,\mathbf{c}), \mathbf{c})}{\partial I} \equiv \frac{w}{p}. (\#eq:eqn2)
\end{equation}

Identity \@ref(eq:eqn2) states that the necessary elements of the information needed for profit maximization are the input price, the output price, and the partial derivative of the yield function with respect to the input. While knowledge of the yield function $f(I, \mathbf{c})$ in Eq. \@ref(eq:eqn1)) is sufficient to solve the maximization problem above (since its derivative can be calculated), identity \@ref(eq:eqn2) shows that it is not a necessary condition; rather, knowing the "change" in yield in response to a change in input rate is.

Recent papers using ML methods to estimate the effectiveness of various agricultural input management strategies first estimate yield function, $f(I, \mathbf{c})$ (e.g., @krause2020random; @barbosa2020modeling; @Gardner21) and then derive changes in yield associated with changes in input rate to estimate site-specific input recommendations. On the other hand, Causal Forest is designed specifically to estimate a discrete version of Eq. \@ref(eq:eqn2) directly, examining the discrete changes in yield that result from discrete *changes* in the input application rate $I$, bypassing the estimation of $f(I, \mathbf{c})$. It may be suspected that CF will outperform other ML methods that identify EOIR indirectly through yield prediction. We now details how CF works below and why CF may perform better in predicting EOIR than other prediction-oriented ML methods used in the previous studies.

## Causal Forest

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CT), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf (please see Appendix for an example of a causal tree).

While both RF and CF build trees by recursively partitioning samples, they differ critically in their criteria for splitting. In RF, the sample is split to minimize the mean squared error (MSE) of yield prediction in building trees. Unfortunately, the MSE of treatment effect predictions is not useful for building a CT since the true treatment effects are never observable. However, @athey2016recursive showed that minimizing the expected MSE of the treatment effect is equivalent to maximizing the variance of treatment effects across the resulting two leaves and minimizing the within-leaf variance. Consequently, by splitting samples in a way that maximizes the variance of treatment effects across the resulting two leaves, CF overcomes the fundamental problem created by the unobservability of true treatment effects. CF requires that the tree-building process is "honest" for treatment effects estimation to be unbiased. Honest tree-building first randomly splits in two the training data samples, then uses one of the subsamples to determine how to split the tree, and uses the other to estimate the treatment effects [@athey2016recursive; @Wager2018a]. This honest subsampling technique avoids overestimation of the heterogeneity of estimated treatment effects. Since CF splits samples to maximize the heterogeneity of estimated treatment effects, using the same sample for both the sample splitting and treatment effect calculations renders itself sensitive to noise or outliers, allowing high noise levels to exaggerate the treatment effects.

## CF vs. Prediction-oriented ML methods

Whether the indirect (yield-prediction-first) and direct approach is used can lead to meaningful statistical differences when it comes to EOIR estimation. The conventional prediction-oriented ML methods are designed to predict yield, $y_i(I,\mathbf{c})$, well, but not necessarily treatment effects. Since the loss function used in prediction-oriented ML uses (typically squared) residuals in yield prediction, prediction-oriented ML is naturally optimized for predicting the "level" of yield instead of identifying the causal impact of the input on yield. CF, on the other hand, is designed specifically to estimate the causal impact of input. Indeed, @Wager2018a used Monte Carlo simulations to show that CF provided more accurate estimates of heterogeneous treatment effects than those provided by RF. However, the underlying data generating process of their simulations was extremely simple, with the treatment effect influenced by only a single variable. CF has not been tested for more complicated data generating processes, such as those represented by yield response functions in which the treatment effect of an input can be affected by multiple variables in a non-linear fashion. Moreover, CF has never been compared to more advanced ML methods such as BRF and CNN that are considered to have prediction capabilities greater than RF's. The study reported here builds on @Wager2018a to compare CF to these prediction-oriented ML methods (not just RF, but also BRF and CNN) in more realistic agronomic experiment settings.

# Materials and Methods

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Then, for each simulation round, one of the fields was used as a training data set, and another field that is different from the one used as a training data set in that round was used as a validation data set. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

## Data Generation

### Simulated Experimental Fields
<!-- experimental field: area: 51ha, plot: 384, subplot:1536, cell:55296   -->

The simulated field was created based on an actual 51-ha Illinois field on which the Data-Intensive Farm Management (DIFM) project [@bullock2019] has run multiple OFPEs. Figure \@ref(fig:field-map) illustrates that simulated trials consisted of $384$ 18.3 m $\times$ 73.2 m "plots," each of which was assigned an N fertilizer application rate. Each plot was made up of four (4-rows $\times$ 1-column) 18.3 m $\times$ 18.3 m "subplots," which were unit of analysis used in subsequent statistical analysis. Each subplot contained of a 6-rows $\times$ 6-columns grid of thirty-six 3.05 m $\times$ 3.05 m "cells." All field characteristics and yield data were generated at the cell level.

```{r field-map, out.width="50%", fig.cap = "Partitioning of field into plots, subplots and cells"}
# knitr::include_graphics("./GitControlled/Writing/field_structure.png")
readPNG(here("GitControlled/Writing/field_structure.png")) %>% grid.raster()
```

### Underlying Yield Response Function

We modeled the yield response function as having the following Mitscherlich-Baule functional form, which is flexible in its shape, allows for the existence of a yield plateau, and has been shown to be a plausible specification for N fertilization models for corn [@frank1990; @llewelyn1997; @paris1992].

\begin{equation}
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) N)) + \varepsilon. (\#eq:eqn5)
\end{equation}

The first term on the right-hand side of Eq. \@ref(eq:eqn5) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are functions of a vector of field characteristics variables $\mathbf{c}=(c_1,...,c_K)$, which vary spatially on the field. For example, $\alpha$ and $ymax$, which affect yield independently from the applied N fertilizer rate, might be interpreted as depending on the ingredient soil N content supplied by residual N, soil organic matter, etc. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$.

A fictional "researcher" is modeled as not knowing the function form of $f$, as having limited knowledge about which variables make up $\mathbf{c}$, or even the number of elements of $\mathbf{c}$, and as having limited ability to measure the values that many of the variables that constitute $\mathbf{c}$ take on across the field.

### Cell-specific Yield Functions and True Economically Optimal N Rates

Let $i \in \{1,2,...,1536\}$ index a simulated field's subplots and $j \in \{1,2,...,36\}$ index cells within subplots. Let $N^{i}$ be the N fertilization rate on plot $i$, and let $\mathbf{c}^{i,j}$ be the value of $\mathbf{c}$ on cell $(i,j)$, Eq. \@ref(eq:eqn6) shows the yield level on cell $j$ in subplot $i$ is affected by the N rate applied on subplot $i$:

\begin{equation}
f(N^{i}, \mathbf{c}^{i,j}) = ymax(\mathbf{c}^{i,j})(1-exp(\alpha(\mathbf{c}^{i,j}) + \beta(\mathbf{c}^{i,j}) N^{i})) + \varepsilon^{i,j}. (\#eq:eqn6)
\end{equation}

To simplify notation, denote  $\alpha(\mathbf{c}^{i,j})$ as $\alpha^{i,j}$, $\beta(\mathbf{c}^{i,j})$ as $\beta^{i,j}$, and $ymax(\mathbf{c}^{i,j})$ as $ymax^{i,j}$. Then, a cell-specific yield function can be written as,

\begin{equation}
f^{i,j}(N) = ymax^{i,j}(1-exp(\alpha^{i,j} + \beta^{i,j} N^{i})) + \varepsilon^{i,j}. (\#eq:eqn7)
\end{equation}

Parameters $\alpha^{i,j}$, $\beta^{i,j}$, $\beta^{i,j}$, and $\varepsilon^{i,j}$ were modeled as spatially autocorrelated using unconditional Gaussian geostatistical simulation based on the spherical variogram model. The `gstat` package (version 2.0.6 [@gstat1; @gstat2]) in R was used for this purpose. Table \@ref(tab:variogram-parameters) shows the main variogram parameter values used in generating $\alpha^{i,j}$, $\beta^{i,j}$, $ymax^{i,j}$, and $\varepsilon^{i,j}$, which were chosen to generate yields consistent with those observed in experiments conducted by the DIFM project. (Codes used to generate parameter values are available at https://github.com/tmieno2/ML_VRA.git.) The range and nugget of all parameters were assumed to be $400$ m and 0. The means of $ymax$, $\alpha$, $\beta$, and $\varepsilon$ were assumed to be $12000$ kg/ha, $-0.5$, $0$ and $0$, and their sills were assumed to be $2.0 \times 10^{6}$ kg/ha, $0.02$, $1.0$ and $0.015$. Figure \@ref(fig:fig-parameters) shows an example of the generated parameters.

<br>

@@

Each cell $(i,j)$'s "true" optimal N application rate, denoted $N_{opt}^{i,j}$, was calculated by solving a profit maximization problem:

\begin{equation}
N_{opt}^{i,j} = \operatorname*{argmax}_{N}(P_{C} f^{i,j}(N) - P_N N), (\#eq:eqn8)
\end{equation}

where, $p =\$0.138/kg$ and $w =\$1.323/kg$ were the assumed prices of the corn and N fertilizer. Figure \@ref(fig:fig-optN) maps an example of the identified "true" cell-specific optimal N rates.

```{r fig-optN, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a spatial distribution of true optimal N rate"}
vis_optN_cell
```

<br>

### Trial Design

In each simulation round, five field-trial N rates were determined based on the range of the round's cell-specific EONRs. 
Let $Q_q$ ($q \in [0,1]$) denote the $100 \times q$ percent quantile of the round's cell-specific EONRs. Then, the five trial design rates were $N_1 = Q_{0.05} - 20$, $N_5 = Q_{0.95} + 20$, $N_2 = N_1 + 0.25(N_5 - N_1)$, $N_3 = N_1 + 0.50(N_5 - N_1)$, $N_4 = N_1 + 0.75(N_5 - N_1)$. 
In each round, a Latin square design was followed to assign each plot one of the five N rates so established. Figure \@ref(fig:fig-nExp) shows an example of a trial design. The N rate was orthogonal to the error term in every simulation's example of trial design. This is important because if they were correlated, bias in the estimation of the impact of N on yield would be introduced into all of the ML methods. Since the treatment effect at each leaf is obtained simply by taking the difference of the average value of the dependent variable between the control and treated group within the leaf, CF requires the exogeneity of the treatment conditional on the independent variables to estimate the treatment effect consistently. Note that the violation of this condition would lead to biased estimation using the CF and the other ML methods.

```{r fig-nExp, fig.width=3, fig.height=4, fig.cap = "An illustrative example of experimental design of N"}
field_Ndesign
```

<br>

### Yield Data Generation and Aggregation

A cell-level data set for the whole field, $\{(y^{i,j}, N_{opt}^{i,j}, N^{i,j}, \mathbf{c}^{i,j}): i \in \{1,2,...,1536\}, j \in \{1,2,...,36\}\}$ was generated. For a generic cell $(i,j)$, yield $y^{i,j}$ was generated through Eq. \@ref(eq:eqn5), applying parameter values $\mathbf{c^{i,j}}=(\alpha^{i,j}, \beta^{i,j}, ymax^{i,j})$, the disturbance term level $\varepsilon^{i,j}$, and the cell's assigned experimental N rate, $N^{i,j}$. Figure \@ref(fig:field-yield-cell) presents the yield map from one of the simulations.

```{r field-yield-cell, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a yield distribution map at the cell resolution"}
vis_yield_cell
```

<br>

Data were aggregated up to the subplot level by taking the mean of the cell-level values inside each subplot, which were the unit of observation. The subplot data were therefore

\begin{equation}
\{(y^{i}, N_{opt}^{i}, N^{i}, \mathbf{c}^{i}) = \frac{1}{36}\sum_{j=1}^{36} (y^{i,j}, N_{opt}^{i,j}, N^{i,j}, \mathbf{c}^{i,j}): i \in \{1,2,...,1536\}\}. (\#eq:eqn9)
\end{equation}

Figure \@ref(fig:field-yield-subplot) shows an example of observed yield data. Note that data generation at the cell level was conducted to reflect the real-world spatial heterogeneity in soil/field characteristics within the subplot level. But data aggregation to the subplot level was conducted because the large sizes of harvesters make it impossible to monitor yields accurately at smaller scales.

```{r field-yield-subplot, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a yield distribution map at the subplot resolution"}
vis_yield_subplot
```

<br>

## Modeling Scenarios

To evaluate the performances of RF, BRF, CNN and CF, we examined four modeling scenarios, each defined by a dataset assumed to be available to the researcher. The first modeling scenario, denoted "aby" assumed that the researcher has a dataset of the actual values $\alpha^{i,j}$, $\beta{i,j}$, $ymax^{i,j}$ for every cell $(i,j)$ and use $\alpha$, $\beta$ and $ymax$ as covariates in the ML models. This scenario represents an ideal situation for the researcher, in which all the relevant variables are observed, but the true functional form of heterogeneous yield response functions are not known. More realistic scenarios can be compared with this ideal scenario.

In subsequent modeling scenarios, we made the simulated experimental data more realistic by varying the degree to which the researcher understood what the needed covariates' values were those in the data set. In the second scenario, denoted "abytt," we introduced additional variables $\theta_1$ and $\theta_2$, and the researcher's data set contained values of $\alpha^{i}$, $\beta^{i}$, $ymax^{i}$, $\theta_1^{i}$ and $\theta_2^{i}$ for all subplots $i=1,...,1536$. $\theta_1$ and $\theta_2$ had no effect on yield, but their cell-level values were spatially correlated with $\beta^{i,j}$, and so their subplot-level values were correlated with $\beta^{i}$. This scenario was meant to reflect the lack of scientific consensus about which variables explain heterogeneous yield response to N, making it likely that the researcher would include irrelevant variables like $\theta_1$ and $\theta_2$ in real-world data analysis.

In the third scenario, denoted "aabbyy," the researcher's data set did not contain values of $\alpha^{i}$, $\beta^{i}$, or $ymax^{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Specifically, for each cell $(i,j)$, we assigned one of a set of spatially autocorrelated weights $r_{\alpha}^{i,j} \in (0,1)$ to the $\alpha$ variable, another set of spatially autocorrelated weights $r_{\beta}^{i,j} \in (0,1)$ to the $\beta$ variable, and another set of spatially autocorrelated weights $r_{ymax}^{i,j} \in (0,1)$ to the ${ymax}^{i,j}$ variable. (There was no spatial correlation between the $r_{\alpha}^{i,j}$ and $r_{\beta}^{i,j}$ data, between $r_{\alpha}^{i,j}$ and $r_{ymax}^{i,j}$ data, or between $r_{\beta}^{i,j}$ and $r_{ymax}^{i,j}$.) Six new covariates were created with the weights and the original covariates $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$. 
These were 
$\alpha_{1}^{i,j} = r_{\alpha}^{i,j} \alpha^{i,j}$, $\alpha_{2}^{i,j} = (1- r_{\alpha}^{i,j}) \alpha^{i,j}$,
$\beta_{1}^{i,j} = r_{\beta}^{i,j} \beta^{i,j}$, $\beta_{2}^{i,j} = (1- r_{\beta}^{i,j}) \beta^{i,j}$,
$ymax_{1}^{i,j} = r_{ymax}^{i,j} ymax^{i,j}$, and $ymax_{2}^{i,j} = (1- r_{ymax}^{i,j}) ymax^{i,j}$.
For each $\alpha$, $\beta$ and $ymax$, these procedures created two new covariates that were highly spatially correlated with the original covariate, which was meant to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax^{i,j}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

The fourth scenario, denoted "aabbyytt," had the same components as aabbyy, but also included the $\theta_1$ and $\theta_2$ covariates.

## Training Models and Estimating Site-specific EONRs

### Training and Estimation of Site-specific EONRs by RF, BRF, and CNN

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF and BRF predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF, BRF, and CF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving honest tree-building process), and the optimal values for these parameters were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same data set was used. To reduce the variance of predictions for prediction accuracy, we set the number of trees (*num.trees*) at a high value, $4000$.

CNN leaves room for researchers to determine its architecture. We used a slightly modified version of one of the multi-stream CNN architectures proposed by @barbosa2020modeling, called "Late Fusion." Briefly stated, explanatory variables except N rate (e.g. $\alpha^{i,j}$, $\beta^{i,j}$, etc.) entered the model as a set of $6 \times 6$ element rasters.The input size of the N rate is treated as $1 \times 1$ because it is spatially homogeneous within each subplot. First, each input was connected to an independent convolutional layer with eight $3 \times 3$ filters each with stride one, followed by a $2 \times 2$ max-pooling layer with stride two. Then, a fully-connected rectified linear unit (ReLU) layer with sixteen neurons was added after each max-pooling layer, followed by a single ReLU neuron. Finally, multiple neurons were concatenated and fed to the fully-connected ReLU layer with sixteen neurons, followed by an output with a linear activation function. @barbosa2020modeling demonstrated that among several CNN architectures, this architecture modeled crop yield response to N rate management best. CNN was implemented in *Python* v3.7.6 using *Pytorch* v1.7.0 [@paszke2017automatic]. The Adam optimizer [@kingma2014adam] was used with a learning rate of $0.001 \%$ (default value). To avoid over-fitting, early stopping was used to monitor validation loss with a ten epochs of patience.

Let $\mathbf{\Omega^{i}}$ denote a list of subplot-level explanatory variables (soil/field characteristics). Further, let $\hat{g}_m(N,\mathbf{\Omega})$ denote an estimated yield response function, estimated by model $m$ ($m$ = RF, BRF, or CNN). Then, yield at $i$ from model $m$ is $\hat{g}_m(N,\mathbf{\Omega^{i}})$. The site-specific EONR for each model can then be found by solving the following profit maximization problem for all $i$ for each of the models:

\begin{equation}
\hat{N}_{opt}^{i} = \operatorname*{argmax}_{N}(p \hat{g}_m(N, \mathbf{\Omega^{i}}) - w N), (\#eq:eqn8)
\end{equation}

where $p$ and $w$ are the prices of corn and N.

### Training and Estimation of Site-specific EONRs by CF

Unlike RF, BRF, and CNN, CF estimates the impact of a binary treatment. In our context, CF estimates changes in yields (i.e., $\hat{\tau}_{N^{con} \rightarrow N^{tre}}(\mathbf{\Omega^{i}})$) caused by changes in N application rates from one experimental rate (control N rate, i.e., $N^{con}$) to another experimental rate (treatment N rate, i.e., $N^{tre}$). Since we have five N application rates, four "experiments" (treatments) are identified based on four ($N^{con}$, $N^{tre}$) combinations. A possible grouping to make such pairwise N application rate combinations, which we call the "CF-base," used $(N_1, N_2),(N_1, N_3),(N_1, N_4)$, and $(N_1, N_5)$. In this approach, the treatment effect is always estimated against the lowest N rate ($N_1$) as $N^{con}$.

For estimating site-specific EONR, the trained CF-base were used to predict site-specific changes in yields resultant from changes in application rates from $N_1$ to $N_m$ $(m\in\{2,3,4,5\})$. Let $\Delta Y_{N_1 \rightarrow N_m}(\mathbf{\Omega^{i}})$ be the the estimated treatment effect of changing subplot $i$'s application rate from $N_1$ to $N_m$ for site $i$.

Since the CF-base approach always estimates the treatment effects using the lowest treatment rate $N_1$, the predicted treatment effects already represent differences in yields the from yield at $N_1$:

>>>>

# Appendix A: Causal Tree Example {#ctree-ex}

Figure \@ref(fig:ct-ex) shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$)($dS/m$) and topographical $\textit{slope}$ ($degree$) as variables to explain the heterogeneity of the impact of the treatment impact.

```{r ct-ex, fig.width=5, fig.height=5, fig.align="center", fig.cap = "An illustrative example of causal tree"}
fancyRpartPlot(opfit, sub = "")
```

NOTE: The first number in each node indicates the predicted treatment effects (in kg/ha). The number and the percentage of observations belonging to the node are shown below that.

In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example. The first data split criterion is whether the $\textit{ecs}$ value is greater or less than $25$. The observations with $\textit{ecs}$ greater than or equal to $25$ move to the left node, while those with $\textit{ecs}$ less than $25$ move to the right. For each of the subsamples, the treatment effect is identified as the difference in the mean yields of the treated and control groups within the subsample. The treatment effect for the low-$\textit{ecs}$ observations was $619$ kg/ha, and for the high-$\textit{ecs}$ observations was $982$ kg/ha. This means that for the impact of the increase in seed rate the treatment effect is heterogeneous over space, having a larger positive impact on yield in high-$\textit{ecs}$ areas. Trees are further developed to have more groups of subsamples to allow for more flexible representations of the heterogeneous treatment effects. The estimated heterogeneous treatments effects can then be used to identify site-specific EOIRs. It is noteworth that at no point CF estimates yield "level".

# Appendix B: Parameter Generation

## Base parameters $\alpha$, $\beta$, and $ymax$ {#par-gen}
Parameters $\alpha^{i,j}$, $\beta^{i,j}$, $ymax^{i,j}$, and $\varepsilon^{i,j}$ were modeled as spatially autocorrelated using unconditional Gaussian geostatistical simulation based on the spherical variogram model. The `gstat` package (version 2.0.6 [@gstat1; @gstat2]) in R was used for this purpose

Table \@ref(tab:variogram-parameters) shows the main variogram parameter values used in generating $\alpha^{i,j}$, $\beta^{i,j}$, $ymax^{i,j}$, and $\varepsilon^{i,j}$, which were chosen to generate yields consistent with those observed in experiments conducted by the DIFM project. (Codes used to generate parameter values are available at https://github.com/tmieno2/ML_VRA.git.) The range and nugget of all parameters were assumed to be $400$ m and 0. The means of $ymax$, $\alpha$, $\beta$, and $\varepsilon$ were assumed to be $12000$ kg/ha, $-0.5$, $0$ and $0$, and their sills were assumed to be $2.0 \times 10^{6}$ kg/ha, $0.02$, $1.0$ and $0.015$. See Figure \@ref(fig:fig-parameters) for an example of the generated parameters.

```{r fig-optN, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a spatial distribution of true optimal N rate"}

vis_optN_cell
```

@@

## Splitting the base parameters {#split-pars}

For scenarios "aabbyy" and "aabbyytt", each of the base parameters $\alpha$, $\beta$, and $ymax$ are splitted into two part in a spatially correlated manner. Speficiallu, for each cell $(i,j)$, we assigned one of a set of spatially autocorrelated weights $r_{\alpha}^{i,j} \in (0,1)$ to the $\alpha$ variable, another set of spatially autocorrelated weights $r_{\beta}^{i,j} \in (0,1)$ to the $\beta$ variable, and another set of spatially autocorrelated weights $r_{ymax}^{i,j} \in (0,1)$ to the ${ymax}^{i,j}$ variable. Each of the weight variables are generated using unconditional Gaussian geostatistical simulation just like the base parameters. (There was no spatial correlation between the $r_{\alpha}^{i,j}$ and $r_{\beta}^{i,j}$ data, between $r_{\alpha}^{i,j}$ and $r_{ymax}^{i,j}$ data, or between $r_{\beta}^{i,j}$ and $r_{ymax}^{i,j}$.) Six new covariates were created with the weights and the original covariates $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$. These were 
$\alpha_{1}^{i,j} = r_{\alpha}^{i,j} \alpha^{i,j}$, $\alpha_{2}^{i,j} = (1- r_{\alpha}^{i,j}) \alpha^{i,j}$,
$\beta_{1}^{i,j} = r_{\beta}^{i,j} \beta^{i,j}$, $\beta_{2}^{i,j} = (1- r_{\beta}^{i,j}) \beta^{i,j}$,
$ymax_{1}^{i,j} = r_{ymax}^{i,j} ymax^{i,j}$, and $ymax_{2}^{i,j} = (1- r_{ymax}^{i,j}) ymax^{i,j}$. The codes used to implement the abovementioned parameter splitting is available at https://github.com/tmieno2/ML_VRA.git.
